---
title: "Lab 6 Comparing Two Means"
author: "Amaya Jerdee"
date: "5/3/2021"
output: github_document
---
Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

```{r}

# load packages -----------------------------------------------------------
library(tidyverse)
# read data ---------------------------------------------------------------

fish <- read_csv("chap12q19ElectricFish.csv")

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

head(fish)

str(fish)
# perform tasks -----------------------------------------------------------

t.test(formula = species ~ location, data = fish_long)  #t-test

ttest_results <- t.test(formula = species ~ location, data = fish_long)

ttest_results

diff(ttest_results$estimate)

```

## Question A

> What is the mean difference in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence intervalof this mean difference. Show your code and write a sentence giving your answer.

ANSWER

The mean difference in the number of species between areas is 1.8334 and the 95% confidence interval is -4.587031 to 8.253697.


```{r}
fish_long_data <-
  fish_long %>% 
  group_by(location) %>% 
  summarize(
    n = n(),
    mean = mean(species),
    sd = sd(species),
    sem = sd/sqrt(n),
    upper = mean + 1.96 * sem,
    lower = mean - 1.96 * sem
  ) %>% 
  print()

fish_long %>% 
  ggplot(aes(x = location, y = species)) + 
  geom_jitter(aes(color = location), 
              shape = 16, size = 3, 
              alpha = 0.3, width = 0.4) + 
  geom_errorbar(aes(y = mean, ymax = upper, ymin = lower), 
                data = fish_long_data, 
                width = 0.1, size = 0.8) + 
  geom_point(aes(y = mean), 
             data = fish_long_data, 
             size = 3) + 
  scale_color_manual(values = c("dark blue","red")) + 
  theme_minimal() + 
  guides(color = "none")
```
## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

According to the t-test there is no signigicant difference between the species and the location due to p>0.05. The difference between the means falls within the 95% confidence interval. With this being said, we fail to reject the null hypothesis. 

```{r}
t.test(formula = species ~ location, data = fish_long)
```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

We assumed that the number of species were normally distributed, the measurements were taken from a random sample, and the standard deviation of the number of species were equivalent in both samples when performing the two-sample t-test. When looking at the graph below, the assumption that the number of species were normally distributed was not accurate. 

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 5, 
    alpha = 0.5, 
    position = "identity"
  ) +
  scale_fill_manual(values = c("darkorange","cyan4")) +
  theme_minimal()
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.



```{r}
Crab <- read_csv("chap15q27FiddlerCrabFans.csv") %>% 
  rename(type = crabType, temp = bodyTemperature) 
Crab

temp_means <- 
  Crab %>% 
  filter(!is.na(temp)) %>%     #remove missing values 
  group_by(type) %>% 
  summarize( 
    mean = mean(temp), 
    sd = sd(temp), 
    n = n(), 
    sem = sd / sqrt(n), 
    upper = mean + 1.96 * sem, 
    lower = mean - 1.96 * sem
  ) %>% 
  print() 
```
## Question D

Graph the distribution of body temperatures for each crab type:

```{r}
ggplot(data = Crab, aes(x = type, y = temp)) + 
  geom_jitter(aes(color = type), 
              width = 0.1, 
              alpha = 0.7, 
              show.legend = FALSE, 
              na.rm = TRUE) + 
  geom_errorbar(aes(y = mean, ymin = lower, ymax = upper), 
                data = temp_means,
                width = 0.1, position = position_nudge(0.3)) + 
  geom_point(aes(y = mean), data = temp_means, 
             position = position_nudge(0.3)) + 
  scale_color_manual(values = c("darkorange","darkorchid","cyan4","blue3"))
```

## Question E
> Does body temperature varies among crab types? State the null and alternative hypothesis, conduct an ANOVA, and interpret the results.

ANSWER 

The null hypothesis is that the mean body temperature is equal between the crab types. The alternative hypothesis is that the mean body temperatureis different in at least one type of crab. The null hypothesis is rejected while the alternative hypothesis is accepted (p<0.05). The difference between the mean body temperature in at least one type of crab and the mean body temperature of the other types of crbas is significantlly statistically different.

```{r}
aov_crab_summary <- 
  aov(temp ~ type, data = Crab) 
aov_crab_summary 

summary(aov_crab_summary)

```

